Modelling Mass Migration based on Spatial\\ Generative Agent-Based Trade Simulation

### 2017-05-26 paper/paper.tex
Minor typos

\subsection{Process overview and scheduling}
Our simulation works with synchronous updates (ticks), meaning all agents do the same things at the same time. The actions in each tick can be divided into three main parts:
\subsubsection{Resource distribution}
	To distribute the resources on the grid, we us the function
	\begin{equation}
		f_\mathit{dist}(\mathit{pos},\, \mathit{time}) \to [0,\, 1]
	\end{equation}
	which takes the position of a parcel as an input and calculates the percentage which determines how much of the total resources this specific parcel receives. To obtain the actual amount of resources distributed, we multiply this percentage with the total need for resources and the $\mgf$:
	\begin{equation}
		resource = f_\mathit{dist}(\mathit{pos},\, \mathit{time}) \cdot n \cdot \col \cdot \mgf
	\end{equation}
	where $n$ is the initial number of agents. Using this formula, we see that the total distribution of resources is equal to the total needs:
	\begin{alignat}{2}
		&\sum_{p \in \mathit{Parcels}} f_\mathit{dist}(\mathit{p.pos},\, \mathit{time}) \cdot n \cdot \col &&= n \cdot \col\\
		\Longleftrightarrow \quad &\sum_{p \in \mathit{Parcels}} f_\mathit{dist}(\mathit{p.pos},\, \mathit{time}) &&= 1
	\end{alignat}

\subsubsection{Trade requests}
	Now all agents have the possibility to propose a trade with exactly one of their neighbours. They do this by finding the best match for a trading partner in the list of their neighbours. The best trading partner is found by evaluating the following formula:
	\begin{equation}
		\argmax_{a \in \mathit{Neighbours}} \left( \gdf \cdot a.\mathit{resource} + (1-\gdf) \cdot \mathit{sympathy}(a) \right)
	\end{equation}
	where $sympathy()$ is the mapping from a neighbour of an agent to a sympathy value.

\subsubsection{Trade transaction}
	 Now all agents can view the list of their neighbours that have proposed a trade with them. The individual willingness-to-trade factor decides with what probability an agent considers accepting a trade in each tick. If an agent decides to accept a trade he finds a best match for a trading partner with the same formula we already used above. A trade causes the sympathy of the giving agent towards the receiving agent to sink by 1, and conversely for the receiving agent towards the giving agent to rise by 2. This mechanism ensures that frequent interactions between a pair of agents results in a strengthened relationship:
	 \begin{equation}
	 	\lim_{n \to \infty} n \cdot (p\cdot 2 - (1 - p)\cdot 1) = \lim_{n \to \infty} n \cdot (3\cdot p - 1) = \begin{cases}
	 	= \infty,\quad &\text{if $p > \frac{1}{3}$}\\
	 		= 0, &\text{if $p = \frac{1}{3}$}\\
	 		= -\infty, &\text{if $p < \frac{1}{3}$}
	 	\end{cases}
	 \end{equation}
	 where $p$ denotes the ratio of mutual trading. This means, that if agent $A$ does not trade more than twice as often with agent $B$ than the other way around, their respective sympathy will increase linearly over time.

\subsubsection{Agent movements}
		After all the trades have been completed the agents will update their position according to the following schema:
\begin{enumerate*}
	\item
		The agent checks whether he has received more resources this tick an in the previous one, if so, he keeps his position, else he moves to his previous position.

@@

	social behaviour:
	\begin{description*}
		\item[$\gdf$:]
			The biggest part of our agent's characters is decided by the value of the gold-dig-factor (cf. gold-dig-factor in the section Entities, state variables and scales). The idea behind this is that a $\gdf$ of 1 causes behaviour similar to socialist/communist ideas while a $\gdf$ of 0 leads to phenomena similar to capitalism. Capitalism favours accumulation of money, be it in the hands of individuals or corporations [Karl Marx, Das Kapital, vol. 1, ch. 25]. On the other side of the spectrum socialism favours the equal distrubution of money as can be seen in socialist ideas such as progressive taxes [Moyes, P. A note on minimally progressive taxation and absolute income inequality]. We mirror this using the $\gdf$ in the following way: an agent with a high $\gdf$ will, when considering all agents that have proposed him a trade, favour the agent with the least resources over any other agent, no matter how much sympathy he has for any of them. It is obvious that this favours the equal distribution of resources since in a scenario of unequally distributed wealth with all agents having a $\gdf$ of 1 all poor agents will propose trades to (really it's requestint resources from) rich agents which, in turn will accept these trades without consideration. On the other hand a $\gdf$ of zero of near zero will favour the emergence of groups of agents which will exclusively help other agents of the group. Poor agents requesting resources will lose the sympathy of richer agents because requesting resources will lead to a decrease of the sympathy factor while rich agents who don't have to request resources very often will maintain a high sympathy factor.

		\item[resource:]
			The one resource of our model can be thought of as being money. Money can take have multiple functions: among others it is a medium of exchange, a measure of value and a store of value [William Stanley Jevons, Money and the Mechanism of Exchange]. In our model the agents need the resource (from now on denoted as money) in order to survive. It acts therefore as a medium of exchange, the agents can, hypothetically, go and buy food and whatever they need in order to survive. At the same time it is a store of value; as we'll see in the next section agents give away money but at the same time they obtain another abstract value, the sympathy of the receiving agent and therefore some form of insurance for the future. The value is not lost but stored in another form. Money also acts as a measure of value, it allows us to quantify the success of individual agents and groups of agents by assigning them and their position a value.

		\item[Trading:]
			As we've seen trading in our model is somehow unintuitive: agents can request a trade and then may or may not receive money without really giving something back. No trading actually takes place, it's more giving than trading. The assumption here is a creditor-debitor relationship. A rich agent will give money to a poor agent but both will remember that the trade has happened in the form of the changed sympathy. While the creditor is less likely to give to the same debitor many times, the debitor remembers his debt towards the creditor and is more likey to give him money in case of a request. This resembles the real world notion of credit in so far as the creditor is willing to wait for the debitors situation to improve before claiming back his money. How exactly this creditor-debitor relationship is to be interpreted depends again on the $\gdf$ of both participants: with a low $\gdf$ the actual effect of a trade is accurately described by the above, with a high $\gdf$ not so much since the selection of a trading partner depends less on the sympathy level. This is no contradiction though since a high $\gdf$ mirrors, as described before, the affinity of an agent to distribute wealth. Apart from the obvious distribution of resources the action of trading is an implicit channel of information between two agents. We will investigate this role of trade a bit more below in the section on the logical connection between agents.

		\item[Sympathy:]
			The total sympathy level is steadily rising as in each trade the net change of sympathy is +1 in contrast of the somewhat more natural approach of saying, the helper's sympathy towards the helped decreases and the helped agent's sympathy towards it's helper increases. Before explaining this we need to take a step back and explain the term. According to the Cambridge Academic Content Dictionary sympathy can mean two things: "a feeling or expression of understanding and caring for someone else who is suffering or has problems that have caused unhappiness" or "a feeling or expression of support and agreement". We use the latter definition for our model although the two are obviously intertwined in our model sympathy is not a reaction to the state of another agent but rather the state of the relationship to this actor. Among other things the sympathy one feels towards someone else is influenced by the proximity to this person, the similarity of one's state to this persons's state and "one's past and vicarious experiences" [Lowenstein, G.; Small, D. A. (2007). "The scarecrow and the tin man: The vicissitudes of human sympathy and caring"]. We express these factors in the way the sympathy factor changes, especially in the fact that the net sympathy is rising. Obviously the proximity between two individuals (thus, also between agents) increases with more interaction and since trade is The interaction in our model it surely should increase the sympathy. On the other hand an agent that has received help has found itself in a situation of need and is thus more likely to being able to relate to other agents being in such a state. 
	\end{description*}

@@

		\item[Phyisical Connection:]
			Agents which are close to each other are not necessarily logically connected by an edge, nevertheless they share a common piece of information in the shape of their parcel. Close neighbours are likely to be situated on the same parcel, thereby receiving the same amount of money and having a similar situation. As in the real world the milieu is somehow related to the wellbeing of its inhabitants. Obviously the physical position of an agent is related to its logical connections. An agent won't, for example, maintain relationships with agents which are very far away for a long time. We ensure this by periodiaclly updating the logical network of an agent to favour closer agents to agents which are further away.

@@

\subsubsection{Learning}
	learning happens collectively: an agent accepting a trade signals to the receiving agent that the position of the giving agent is superior to his. The movement the receiving agent then takes is an attempt to improve his own position on the grid. This way trading acts as a kind of communication between agents as it does in the real world. Money flowing from urban to rural areas for example often leads to a flow of people from rural to urban areas [Davis, Kingsley. "The Urbanization of the Human Population"].

@@

\subsubsection{Stochasticity}
	age of the agents is modeled with stochasticity: instead of giving each agent an explicit age and fixing the age of death or increasing the chance of death with increasing age we simply set the chance of death for all agents in a certain tick constant to 1/1000. This gives an expected age of 1000 ticks for each agent. Analogously we deal with spawning new agents (giving birth). Similarly we model the behavioural trait of agentswanting to trade or not, the willingness-to-trade factor.

>>>>

\subsection{Process overview and scheduling}
Our simulation works with synchronous updates (ticks), meaning all agents do the same things at the same time. The actions in each tick can be divided into four main parts:

\subsubsection{Resource distribution}
	To distribute the resources on the grid, we us the function
	\begin{equation}
		f_\mathit{dist}(\mathit{pos},\, \mathit{time}) \to [0,\, 1]
	\end{equation}
	which takes the position of a parcel as an input and calculates the percentage which determines how much of the total resources this specific parcel receives. To obtain the actual amount of resources distributed, we multiply this percentage with the total need for resources and the $\mgf$:
	\begin{equation}
		resource = f_\mathit{dist}(\mathit{pos},\, \mathit{time}) \cdot n \cdot \col \cdot \mgf
	\end{equation}
	where $n$ is the initial number of agents. Using this formula, we see that the total distribution of resources is equal to the total needs:
	\begin{alignat}{2}
		&\sum_{p \in \mathit{Parcels}} f_\mathit{dist}(\mathit{p.pos},\, \mathit{time}) \cdot n \cdot \col &&= n \cdot \col\\
		\Longleftrightarrow \quad &\sum_{p \in \mathit{Parcels}} f_\mathit{dist}(\mathit{p.pos},\, \mathit{time}) &&= 1
	\end{alignat}

\subsubsection{Trade requests}
	Now all agents have the possibility to propose a trade with exactly one of their neighbours. They do this by finding the best match for a trading partner in the list of their neighbours. The best trading partner is found by evaluating the following formula:
	\begin{equation}\label{eq: trade_req}
		\argmax_{a \in \mathit{Neighbours}} \left( \gdf \cdot a.\mathit{resource} + (1-\gdf) \cdot \mathit{sympathy}(a) \right)
	\end{equation}
	where $sympathy()$ is the mapping from a neighbour of an agent to a sympathy value.

\subsubsection{Trade transactions}
	 Now all agents can view the list of their neighbours that have proposed a trade with them. The individual willingness-to-trade factor decides with what probability an agent considers accepting a trade in each tick. If an agent decides to accept a trade he finds a best match for a trading partner with the following formula (note the similarities to equation \ref{eq: trade_req}):
	 \begin{equation}
		 \argmin_{a \in \mathit{Neighbours}} \left( \gdf \cdot a.\mathit{resource} - (1-\gdf) \cdot \mathit{sympathy}(a) \right)
	 \end{equation}
	 A trade causes the sympathy of the giving agent towards the receiving agent to sink by 1, and conversely for the receiving agent towards the giving agent to rise by 2. This mechanism ensures that frequent interactions between a pair of agents results in a strengthened relationship:
	 \begin{equation}
	 	\lim_{n \to \infty} n \cdot (p\cdot 2 - (1 - p)\cdot 1) = \lim_{n \to \infty} n \cdot (3\cdot p - 1) = \begin{cases}
	 	= \infty,\quad &\text{if $p > \frac{1}{3}$}\\
	 		= 0, &\text{if $p = \frac{1}{3}$}\\
	 		= -\infty, &\text{if $p < \frac{1}{3}$}
	 	\end{cases}
	 \end{equation}
	 where $p$ denotes the ratio of mutual trading. This means, that if agent $A$ does not trade more than twice as often with agent $B$ than the other way around, their respective sympathy will increase linearly over time.

\subsubsection{Agent movements}
		After all the trades have been completed the agents will update their position according to the following schema:
\begin{enumerate*}
	\item
		The agent checks whether he has received more resources this tick than in the previous one, if so, he keeps his position, else he moves to his previous position.

@@

	social behaviour:
	\begin{description*}
		\item[$\boldsymbol{\gdf}$:]
			The biggest part of our agent's characters is decided by the value of the gold-dig-factor (cf. gold-dig-factor in the section Entities, state variables and scales). The idea behind this is that a $\gdf$ of 1 causes behaviour similar to socialist/communist ideas while a $\gdf$ of 0 leads to phenomena similar to capitalism. Capitalism favours accumulation of money, be it in the hands of individuals or corporations \footnote{Karl Marx. "Das Kapital", vol. 1, ch. 25}. On the other side of the spectrum socialism favours the equal distribution of money as can be seen in socialist ideas such as progressive taxes \footnote{Moyes, P. "A note on minimally progressive taxation and absolute income inequality"}. We mirror this using the $\gdf$ in the following way: an agent with a high $\gdf$ will, when considering all agents that have proposed him a trade, favour the agent with the least resources over any other agent, no matter how much sympathy he has for any of them. It is obvious that this favours the equal distribution of resources since in a scenario of unequally distributed wealth with all agents having a $\gdf$ of 1 all poor agents will propose trades to (really it's requesting resources from) rich agents which, in turn will accept these trades without consideration. On the other hand a $\gdf$ close to zero will favour the emergence of groups of agents which will exclusively help other agents of the group. Poor agents requesting resources will lose the sympathy of richer agents because requesting resources will lead to a decrease of the sympathy factor while rich agents who don't have to request resources very often will maintain a high sympathy factor.

		\item[resource:]
			The one resource of our model can be thought of as being money. Money can take have multiple functions: among others it is a medium of exchange, a measure of value and a store of value\footnote{William Stanley Jevons. "Money and the Mechanism of Exchange"}. In our model the agents need the resource (from now on denoted as money) in order to survive. It acts therefore as a medium of exchange, the agents can, hypothetically, go and buy food and whatever they need in order to survive. At the same time it is a store of value; as we'll see in the next section agents give away money but at the same time they obtain another abstract value, the sympathy of the receiving agent and therefore some form of insurance for the future. The value is not lost but stored in another form. Money also acts as a measure of value, it allows us to quantify the success of individual agents and groups of agents by assigning them and their position a value.

		\item[Trading:]
			As we've seen trading in our model is somehow unintuitive: agents can request a trade and then may or may not receive money without really giving something back. No trading actually takes place, it's more giving than trading. The assumption here is a creditor-debitor relationship. A rich agent will give money to a poor agent but both will remember that the trade has happened in the form of the changed sympathy. While the creditor is less likely to give to the same debitor many times, the debitor remembers his debt towards the creditor and is more likely to give him money in case of a request. This resembles the real world notion of credit in so far as the creditor is willing to wait for the debitors situation to improve before claiming back his money. How exactly this creditor-debitor relationship is to be interpreted depends again on the $\gdf$ of both participants: with a low $\gdf$ the actual effect of a trade is accurately described by the above, with a high $\gdf$ not so much since the selection of a trading partner depends less on the sympathy level. This is no contradiction though since a high $\gdf$ mirrors, as described before, the affinity of an agent to distribute wealth. Apart from the obvious distribution of resources the action of trading is an implicit channel of information between two agents. We will investigate this role of trade a bit more below in the section on the logical connection between agents.

		\item[Sympathy:]
			The total sympathy level is steadily rising as in each trade the net change of sympathy is +1 in contrast of the somewhat more natural approach of saying, the helper's sympathy towards the helped decreases and the helped agent's sympathy towards it's helper increases. Before explaining this we need to take a step back and explain the term. According to the Cambridge Academic Content Dictionary sympathy can mean two things: "a feeling or expression of understanding and caring for someone else who is suffering or has problems that have caused unhappiness" or "a feeling or expression of support and agreement". We use the latter definition for our model although the two are obviously intertwined in our model sympathy is not a reaction to the state of another agent but rather the state of the relationship to this actor. Among other things the sympathy one feels towards someone else is influenced by the proximity to this person, the similarity of one's state to this persons's state and "one's past and vicarious experiences" \footnote{Lowenstein, G. "The scarecrow and the tin man: The vicissitudes of human sympathy and caring", (2007)}. We express these factors in the way the sympathy factor changes, especially in the fact that the net sympathy is rising. Obviously the proximity between two individuals (thus, also between agents) increases with more interaction and since trade is The interaction in our model it surely should increase the sympathy. On the other hand an agent that has received help has found itself in a situation of need and is thus more likely to being able to relate to other agents being in such a state. 
	\end{description*}

@@

		\item[Phyisical Connection:]
			Agents which are close to each other are not necessarily logically connected by an edge, nevertheless they share a common piece of information in the shape of their parcel. Close neighbours are likely to be situated on the same parcel, thereby receiving the same amount of resources and having a similar situation. As in the real world the milieu is somehow related to the well-being of its inhabitants. Obviously the physical position of an agent is related to its logical connections. An agent won't, for example, maintain relationships with agents which are very far away for a long time. We ensure this by periodically updating the logical network of an agent to favour closer agents to agents which are further away.

@@

\subsubsection{Learning}
	learning happens collectively: an agent accepting a trade signals to the receiving agent that the position of the giving agent is superior to his. The movement the receiving agent then takes is an attempt to improve his own position on the grid. This way trading acts as a kind of communication between agents as it does in the real world. Money flowing from urban to rural areas for example often leads to a flow of people from rural to urban areas \footnote{Davis, Kingsley. "The Urbanization of the Human Population"}.

@@

\subsubsection{Stochasticity}
	age of the agents is modeled with stochasticity: instead of giving each agent an explicit age and fixing the age of death or increasing the chance of death with increasing age we simply set the chance of death for all agents in a certain tick constant to 1/1000. This gives an expected age of 1000 ticks for each agent. Analogously we deal with spawning new agents (giving birth). Similarly we model the behavioural trait of agents wanting to trade or not, the willingness-to-trade factor.
